---
title: "Exploratory Data Analysis"
author: "Yashas BVN"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(dplyr)
```

### Exercise 1

```{r, message=FALSE, warning=FALSE}

GameFun <- read_xlsx('GameFun.xlsx')
summary(GameFun)

#Splitting the test and control groups into data frames
GameFun_test <- GameFun %>% filter(test == 1)
GameFun_control <- GameFun %>% filter(test == 0)
```

#### Question 1.1.a

```{r, message=FALSE, warning=FALSE}

#Comparing averages of income, gender and gamer variables in the test and control groups

#income
income_avg_test <- mean(GameFun_test$income)
income_avg_control <- mean(GameFun_control$income)

income_avg_difference <- (income_avg_test - income_avg_control)/income_avg_control

#gender
gender_avg_test <- mean(GameFun_test$gender)
gender_avg_control <- mean(GameFun_control$gender)

gender_avg_difference <- (gender_avg_test - gender_avg_control)/gender_avg_control

#gamer
gamer_avg_test <- mean(GameFun_test$gamer)
gamer_avg_control <- mean(GameFun_control$gamer)

gamer_avg_difference <- (gamer_avg_test - gamer_avg_control)/gamer_avg_control

paste0("The % difference in average for Income, Gender and Gamer variables between test and control is ",
      round(income_avg_difference*100,4),"%, ",
      round(gender_avg_difference*100,4),"%, ",
      round(gamer_avg_difference*100,4),"%")
```

#### Question 1.1.b

As we see, the difference between the averages for the three variables is very low - indicating that the test and control populations are probabilistically equivalent in these variables.

#### Question 1.1.c

If we had run this analysis before executing an experiment and found a large difference between the test and control groups, the following steps need to be taken:

- Regenerate the test and control groups using randomization
- Test again and check if there still exists a difference between the two populations
- If difference still exists, A/B with matching needs to be performed
- This can be done by matching test and control population based on their similarities and assigning them randomly into the groups

#### Question 1.1.d

In cases where there are a large number of samples, if the classical statistical significance test would not work due to the computational requirements, we can use bootstrapping techniques and take smaller samples from this data, take the means for the test and control groups separately for each of the sub-sample. After this, we can get a distribution of these means and perform a classical statistical test to find if these are statistically different.

This technique is also known as Bag-of-little bootstrap.

#### Question 1.2

Calculating the average purchase rates in the test and control group for different sets of users.

```{r, message=FALSE, warning=FALSE}

# a. Comparison 1: All customers
## We will perform a z-test of proportions 

prop_test_1 <- prop.test(x = c(sum(GameFun_test$purchase), sum(GameFun_control$purchase)),
          n = c(nrow(GameFun_test), nrow(GameFun_control)))

paste("For all customers, The average purchase rate in test and control groups are",
      round(prop_test_1$estimate[1]*100,3),"% and", round(prop_test_1$estimate[2]*100,3),
      "% respectively and the absolute difference in the rates is",
      round((prop_test_1$estimate[1]-prop_test_1$estimate[2])*100,3),"%")


# b. Comparison 2: Male and Female customers
## We will perform a z-test of proportions 

#Splitting populations into male and female
GameFun_test_male <- GameFun_test %>% filter(gender == 1)
GameFun_test_female <- GameFun_test %>% filter(gender == 0)

GameFun_control_male <- GameFun_control %>% filter(gender == 1)
GameFun_control_female <- GameFun_control %>% filter(gender == 0)

prop_test_2_male <- prop.test(x = c(sum(GameFun_test_male$purchase), sum(GameFun_control_male$purchase)),
          n = c(nrow(GameFun_test_male), nrow(GameFun_control_male)))

paste("For males, The average purchase rate in test and control groups are",
      round(prop_test_2_male$estimate[1]*100,3),"% and", round(prop_test_2_male$estimate[2]*100,3),
      "% respectively and the absolute difference in the rates is",
      round((prop_test_2_male$estimate[1]-prop_test_2_male$estimate[2])*100,3),"%")

prop_test_2_female <- prop.test(x = c(sum(GameFun_test_female$purchase), sum(GameFun_control_female$purchase)),
          n = c(nrow(GameFun_test_female), nrow(GameFun_control_female)))

paste("For females, The average purchase rate in test and control groups are",
      round(prop_test_2_female$estimate[1]*100,3),"% and", round(prop_test_2_female$estimate[2]*100,3),
      "% respectively and the absolute difference in the rates is",
      round((prop_test_2_female$estimate[1]-prop_test_2_female$estimate[2])*100,3),"%")

# c. Comparison 3: Gamers and Non Gamers
## We will perform a z-test of proportions 

#Splitting populations into gamers and non_gamers
GameFun_test_gamers <- GameFun_test %>% filter(gamer == 1)
GameFun_test_non_gamers <- GameFun_test %>% filter(gamer == 0)

GameFun_control_gamers <- GameFun_control %>% filter(gamer == 1)
GameFun_control_non_gamers <- GameFun_control %>% filter(gamer == 0)

prop_test_2_gamers <- prop.test(x = c(sum(GameFun_test_gamers$purchase), sum(GameFun_control_gamers$purchase)),
          n = c(nrow(GameFun_test_gamers), nrow(GameFun_control_gamers)))

paste("For gamers, The average purchase rate in test and control groups are",
      round(prop_test_2_gamers$estimate[1]*100,3),"% and", round(prop_test_2_gamers$estimate[2]*100,3),
      "% respectively and the absolute difference in the rates is",
      round((prop_test_2_gamers$estimate[1]-prop_test_2_gamers$estimate[2])*100,3),"%")

prop_test_2_non_gamers <- prop.test(x = c(sum(GameFun_test_non_gamers$purchase), sum(GameFun_control_non_gamers$purchase)),
          n = c(nrow(GameFun_test_non_gamers), nrow(GameFun_control_non_gamers)))

paste("For non-gamers, The average purchase rate in test and control groups are",
      round(prop_test_2_non_gamers$estimate[1]*100,3),"% and", round(prop_test_2_non_gamers$estimate[2]*100,3),
      "% respectively and the absolute difference in the rates is",
      round((prop_test_2_non_gamers$estimate[1]-prop_test_2_non_gamers$estimate[2])*100,3),"%")

# d. Comparison 4: Male Gamers and Female Gamers
## We will perform a z-test of proportions 

#Splitting populations into gamers and non_gamers
GameFun_test_gamer_male <- GameFun_test %>% filter(gamer == 1 & gender == 1)
GameFun_test_gamer_female <- GameFun_test %>% filter(gamer == 1 & gender == 0)

GameFun_control_gamer_male <- GameFun_control %>% filter(gamer == 1 & gender == 1)
GameFun_control_gamer_female <- GameFun_control %>% filter(gamer == 1 & gender == 0)

prop_test_3_gamer_male <- prop.test(x = c(sum(GameFun_test_gamer_male$purchase), sum(GameFun_control_gamer_male$purchase)),
          n = c(nrow(GameFun_test_gamer_male), nrow(GameFun_control_gamer_male)))

paste("For gamer males, The average purchase rate in test and control groups are",
      round(prop_test_3_gamer_male$estimate[1]*100,3),"% and", round(prop_test_3_gamer_male$estimate[2]*100,3),
      "% respectively and the absolute difference in the rates is",
      round((prop_test_3_gamer_male$estimate[1]-prop_test_3_gamer_male$estimate[2])*100,3),"%")

prop_test_3_gamer_female <- prop.test(x = c(sum(GameFun_test_gamer_female$purchase), sum(GameFun_control_gamer_female$purchase)),
          n = c(nrow(GameFun_test_gamer_female), nrow(GameFun_control_gamer_female)))

paste("For gamer females, The average purchase rate in test and control groups are",
      round(prop_test_3_gamer_female$estimate[1]*100,3),"% and", round(prop_test_3_gamer_female$estimate[2]*100,3),
      "% respectively and the absolute difference in the rates is",
      round((prop_test_3_gamer_female$estimate[1]-prop_test_3_gamer_female$estimate[2])*100,3),"%")

```

#### Question 1.3

As we see from the case, revenue per acquisition is $37.5. Expected revenue will be calculated as revenue per acquisition x proportion of customers who were acquired.

We then do the test of difference in proportion to get the difference.

```{r, message=FALSE, warning=FALSE}

#a. Comparison 1 - All customers

expected_revenue_test <- as.numeric(prop_test_1$estimate[1])*37.5

expected_revenue_control <- as.numeric(prop_test_1$estimate[2])*37.5

paste("The expected revenue per person in the test group is $", round(expected_revenue_test,3),
      "and in the control group is $", round(expected_revenue_control,3))

#b. Comparison 2 - Female Gamers vs Male Gamers

#Male
expected_revenue_test_male_gamers <- as.numeric(prop_test_3_gamer_male$estimate[1])*37.5

expected_revenue_control_male_gamers <- as.numeric(prop_test_3_gamer_male$estimate[2])*37.5

paste("The expected revenue per person in the test group of male gamers is $", round(expected_revenue_test_male_gamers,3),
      "and in the control group is $", round(expected_revenue_control_male_gamers,3))

#Female
expected_revenue_test_female_gamers <- as.numeric(prop_test_3_gamer_female$estimate[1])*37.5

expected_revenue_control_female_gamers <- as.numeric(prop_test_3_gamer_female$estimate[2])*37.5

paste("The expected revenue per person in the test group of female gamers is $", round(expected_revenue_test_female_gamers,3),
      "and in the control group is $", round(expected_revenue_control_female_gamers,3))

```

As we see from both these groups, the test group are performing much better than the control group in terms of revenue per person.

#### Question 1.4

As we see from the above numbers, in most cases, the test group out-performs the control group in terms of acquisitions and revenue per person.This indicates that Game-Fun should definitely run this promotion again in the future. 

Moreover, there seems to be a larger difference between the test and control groups for a few targeted segments compared to in general. As we see, the difference is the highest between female gamers when compared to all other slices we checked for.

Also, we see that there might not be much of an impact for non-gamers as the purchase rates are almost equal for the test and control groups.

Hence, the most recommended targeted segment for future ad campaigns should be Female gamers. But as the gamer community is overall showing a high difference, the higher target should be all gamers, before diving into the female gamer community. 

### Exercise 2

```{r, message=FALSE, warning=FALSE}

shots <- read.csv('sommer_deger.csv')
summary(shots)
```

Looking into the approaches of the 4 data scientists.

#### Question 2.1.a

```{r, message=FALSE, warning=FALSE}

#Filtering for babies whose mothers were offered Vitamin A shots for their babies
vitamin_a_offered <- shots %>% filter(instrument == 1)

prop_1a <- (vitamin_a_offered %>% summarise(babies_died = sum(outcome)))/nrow(vitamin_a_offered)

paste(round(prop_1a*100,3), "% of babies whose mothers were offered vitamin A shots for their babies died")

```

#### Question 2.1.b

```{r, message=FALSE, warning=FALSE}

#Filtering for babies whose mothers were not offered Vitamin A shots for their babies
vitamin_a_not_offered <- shots %>% filter(instrument == 0)

prop_1b <- (vitamin_a_not_offered %>% summarise(babies_died = sum(outcome)))/nrow(vitamin_a_not_offered)

paste(round(prop_1b*100,3), "% of babies whose mothers were not offered vitamin A shots for their babies died")

```
#### Question 2.1.c

```{r, message=FALSE, warning=FALSE}

mortality_diff_1 <- prop_1a - prop_1b

paste("The difference in mortality is", round(mortality_diff_1*100,3),"%")
```

The assumption we make in this whole experiment (not just this part) is that the group who were offered and those who were not offered are probabilistically equivalent.Also, there would be an assumption of SUTVA (Stable unit treatment value assumption) which assumes that there is no interaction within the units (mothers and babies) which can affect the results of the experiment.

For the difference between these two percentages to be a valid estimate of the causal impact of receiving vitamin A shots on survival, we need to assume that all mothers who were offered the shot actually let their babies receive the treatment. 

As we see from actual data, this assumption does not actually hold true in this case as there are mothers who did not let their babies get treated despite being offered to take the shot. This will make the estimate biased.

#### Question 2.2.a

```{r, message=FALSE, warning=FALSE}

#Filtering for babies who received vitamin A shots
vitamin_a_received <- shots %>% filter(treatment == 1)

prop_2a <- (vitamin_a_received %>% summarise(babies_died = sum(outcome)))/nrow(vitamin_a_received)

paste(round(prop_2a*100,3), "% of babies who received vitamin A died")

```
#### Question 2.2.b

```{r, message=FALSE, warning=FALSE}

#Filtering for babies who received vitamin A shots
vitamin_a_not_received <- shots %>% filter(treatment == 0)

prop_2b <- (vitamin_a_not_received %>% summarise(babies_died = sum(outcome)))/nrow(vitamin_a_not_received)

paste(round(prop_2b*100,3), "% of babies who did not receive vitamin A died")

```

#### Question 2.2.c

```{r, message=FALSE, warning=FALSE}

mortality_diff_2 <- prop_2a - prop_2b

paste("The difference in mortality is", round(mortality_diff_2*100,3),"%")
```
For the difference between these two percentages to be a valid estimate of the causal impact of receiving vitamin A shots on survival, we need to assume that all the people who did not receive the shot were also offered a shot.

In this case, we can see that only about half the total population were offered a shot. Most of the babies who did not get a vitamin A shot were not even offered one.

#### Question 2.3.a

```{r, message=FALSE, warning=FALSE}

#Filtering for babies who were offered and received vitamin A shots
vitamin_a_offered_received <- shots %>% filter(treatment == 1 & instrument == 1)

prop_3a <- (vitamin_a_offered_received %>% summarise(babies_died = sum(outcome)))/nrow(vitamin_a_offered_received)

paste(round(prop_3a*100,3), "% of babies who were offered and received vitamin A died")

```

#### Question 2.3.b

```{r, message=FALSE, warning=FALSE}

#Filtering for babies who were offered and did not receive vitamin A shots
vitamin_a_offered_not_received <- shots %>% filter(treatment == 0 & instrument == 1)

prop_3b <- (vitamin_a_offered_not_received %>% summarise(babies_died = sum(outcome)))/nrow(vitamin_a_offered_not_received)

paste(round(prop_3b*100,3), "% of babies who were offered and did not receive vitamin A died")

```

#### Question 2.3.c

```{r, message=FALSE, warning=FALSE}

mortality_diff_3 <- prop_3a - prop_3b

paste("The difference in mortality is", round(mortality_diff_3*100,3),"%")
```
For the difference between these two percentages to be a valid estimate of the causal impact of receiving vitamin A shots on survival, we need to assume that the population of mothers who let their babies take the shot were equivalent to the population of mothers who did not give a shot to their babies.


#### Question 2.4.a

```{r, message=FALSE, warning=FALSE}

#Computing Wald estimate for the given data set
prop_offered_received <- (vitamin_a_offered %>% filter(treatment == 1) %>% summarise(received = sum(treatment)))/nrow(vitamin_a_offered)

wald_estimate_4a <- (prop_1a - prop_1b)/prop_offered_received

paste("The Wald estimate for the given dataset is", round(wald_estimate_4a,3))

```

#### Question 2.4.b

The Wald estimate is a valid estimate of the causal impact of vitamin A shots on survival, with the assumption that there is a relationship (not necessarily one-to-one) between being offered the shot and actually receiving it. If there is no relationship between these values, it can lead to a biased and inconsistent estimator, and also lead to a 0 value in the denominator - which could be invalid.

Also, we assume that the only relationship between outcome and being offered a vaccine is only through whether the baby received the vaccine or not and not any other relationship.

We also assume that outcome and being offered are independent - that there is no confounding effect.

#### Question 2.4.c

In the recommendation by the first data scientist, the instrumental variable is whether the mother was offered the shot.

```{r, message=FALSE, warning=FALSE}

#Finding the intent-to-treat estimate recommended by first data scientist
outcome_offered <- lm(shots$outcome~shots$instrument)
summary(outcome_offered)

#Finding Wald estimate for the recommendation of fourth data scientist

#Step 1 of the Two stage least square estimate
treatment_instrumental <- lm(shots$treatment~shots$instrument)
summary(treatment_instrumental)

#Getting the estimates of the instrumental variable
new_treatment <- predict(treatment_instrumental)

#Getting the Wald estimate - step 2 
wald_estimator_lm <- lm(shots$outcome~new_treatment)
summary(wald_estimator_lm)

```
As we see, the standard error for intent-to-treat estimate recommended by the first data scientist is 0.0009229 and that of the Wald estimate recommended by the fourth data scientist is 0.0011536.

i. As we see, the standard error of the Wald estimator is larger. From the formula that we used to derive these, we can see that the Wald estimator does not count for the initial randomization that we had performed where as the intent-to-treat estimate takes into consideration the initial randomization of the population.

ii. The standard error for intent-to-treat assessment might be biased because this assessment does not actually considers only the initial randomly assigned offers and does not consider the ones who chose not to get treated. However, it prevents attrition bias as there are no units who are lost from the study. 
On the other hand, the Wald estimate's standard error can be biased as in the step 1 of the 2 stage least squares, the R-squared value of effect of being offered a shot on the treatment is not very good. This can lead to unexplained variation in the outcome that maybe caused by certain unobserved factors that affect both the outcome as well as the treatment.

### Exercise 3

My opinion on the the paper “The Design Versus the Analysis of Observational Studies for Causal Effects: Parallels With the Design of Randomized Trials” by Donald Rubin.

- It talks about RCM - Rubin Causal Model for observational studies
- The paper stresses on importance of designing the study, regardless of whether it is experiment or observational,carefully before making causal inferences from the data. I really like how the author proposes an alternate title for the whole paper 'Design trumps analysis', further stressing on the importance of the study design
- The paper then goes on to define units, treatments and potential outcomes at a future point in time. The key takeaway here is that all this should be defined before looking at an outcome variable. 
- In section 4 of the paper, the author talks about the case of US tobacco litigation that tries to compare healthcare cost and disease rates of smokers and non-smokers who are very similar in other characteristics. What fascinated me is the amount of background variables that were used to define these sets of people. Although this is a common practice, this is the first time I actually got a chance to look at the variables of a study. Also, the author matches the propensity scores of the background covariates to get similar distributions of people between the treatment and control groups.
- Lastly, I liked how there is a Q&A at the end of the paper with questions simulating a trial.
